---
title: "Assignment 5"
author: "Colin Eberl-Coe"
date: "8 March 2016"
header-includes:
  - \usepackage{bm}
output: pdf_document
---
***
```{r, include=FALSE}
rm(list=ls())
load(paste(getwd(),"/.RData",sep=""))
list.of.packages <- c("sp", "dplyr", "stringr","spatstat","fields","gstat","lattice","ggplot2","knitr","rmarkdown")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, library, character.only = TRUE)
select <- dplyr::select
rm(list.of.packages,new.packages)
if(!exists("wipp",.GlobalEnv)){
  W <- read.table(paste(getwd(),"/WIPP_transmissivity.txt",sep=""), header=TRUE, quote="\"")
  wipp <- W %>% filter(logT != min(logT))
  rm(W)
}
```
### 1. Fit the spherical svgm w/o a nugget using `gstat` and overlay the fitted model on a plot of the empirical svgm. {.tabset} 
```{r Spherical model on emp svgm, echo=F, fig.align='center', fig.height=4, fig.width=6}
if(class(wipp) != "SpatialPointsDataFrame"){
  sp::coordinates(wipp) = ~x+y
}
vg <- gstat::variogram(logT~1,data=wipp)
initial <- gstat::vgm(psill=2.4,model="Sph",range=11)
fit.vg <- gstat::fit.variogram(vg, initial, fit.method=2)

# Functions to facilitate plotting
spherical=function(t2,t3,h){
  less.than=which(h<=t3)
  more.than=which(h>t3)
  gam <- rep(NA,length(h))
  gam[less.than] <- t2*(1.5*h[less.than]/t3-0.5*(h[less.than]/t3)^3)
  gam[more.than] <- (t2)*rep(1,length(more.than))
  return(gam)
}
sph_points <- function(x){
  spherical(t2=fit.vg$psill,t3=fit.vg$range,h=x)
}
# Creating dataframe to be passed in to ggplot
df <- vg %>% select(2,3) %>% as.data.frame %>% mutate(g = sph_points(dist)) %>% rename(distance = dist, gam.hat = gamma, gam.sph = g)

# Ggplot object
gp <- ggplot(df, aes(x=distance,y=gam.hat)) + 
  theme_classic()+
  geom_point(size=3) + stat_function(colour="red",fun = sph_points) +
  ggtitle(label = "Empirical Svgm with Overlaid Spherical (No Nugget) Fit") +
  ylab(expression(hat(gamma)))

# Plot the object
gp

```

  * Krige at all of the observed locations. Show that kriging "honors the data". 
```{r Kriging honors, echo=TRUE}
k <- krige(logT~1,locations=wipp,newdata=wipp,model=fit.vg,debug.level=0)
z <- wipp$logT
z.hat <- k$var1.pred
all.equal(z, z.hat)
```
When predicting at the observed values, kriging returns the observed values themselves; thus the actual `logT` values in the dataset and the output of `krige` are identical. This is \textsf{R}'s way of saying that $\hat{Z}(\bm{s}_i) = Z(\bm{s}_i) \ \forall i$.

  * Krige at grid nodes, plotting predictions and kriging variance. 
```{r Gridded Plots, echo=FALSE,fig.show='hold',fig.align='center', fig.height=4, fig.width=6}
x.grid=seq(min(wipp$x),max(wipp$x),len=100)
y.grid=seq(min(wipp$y),max(wipp$y),len=100)
GRID=expand.grid(x.grid,y.grid,KEEP.OUT.ATTRS = F) %>% `colnames<-`(c("gx","gy"))
coordinates(GRID) = ~gx+gy

kg <- krige(logT~1,locations=wipp,newdata=GRID,model=fit.vg,debug.level=0)
par(mfrow=c(1,2))
quilt.plot(x=as.matrix(as.data.frame(GRID)), y = kg$var1.pred,xlab="x",ylab="y")
title("Gridded Predictions")

quilt.plot(x=as.matrix(as.data.frame(GRID)), y = kg$var1.var,xlab="x")
title("Variability of Predictions")
```
    
  * The RMSE for the transmissivity data is 
    
```{r RMSE, echo=FALSE}
loocv_seq <- seq(1:40)
cv <- function(x, sdata){
  if(class(sdata) != "SpatialPointsDataFrame"){
    sp::coordinates(sdata) = ~x+y
  }
  t <- sdata[-x,]
  vg <- gstat::variogram(logT~1,data=t)
  fit.vg <- gstat::fit.variogram(vg, initial, fit.method=2)
  kcv <- krige(formula = logT~1,locations = t,newdata=sdata[x,],model=fit.vg,debug.level=0)
  return(kcv$var1.pred)
}
loocv_resids <- wipp$logT - sapply(loocv_seq,FUN = cv,sdata=wipp)
(rmse.cv <- sqrt( mean( loocv_resids^2 ) ) )
```
    
  * Plot the predicted vs. observed values. Identify the worst prediction 
    
```{r Worst Prediction Plot, echo=FALSE}
par(mfrow=c(1,1))
quilt.plot(x = as.matrix(as.data.frame(wipp)), y=loocv_resids,main="Residuals of LOOCV Kriging\n With Worst Prediction Circled",xlab="x",ylab="y")
points(wipp[which.max(loocv_resids),1],wipp[which.max(loocv_resids),2],pch=1,cex=5)
```

Observation 35 is predicted the worst, with a residual of 1.96.

### 2.  Inverse Distance Weighting{.tabset} 
  * Does IDW honor the data as Kriging does? 
```{r Does IDW honor data?, echo=TRUE}
invdw <- idw(logT~1,locations=wipp,newdata=wipp,idp=p,debug.level=0)
z <- wipp$logT
z.hat <- k$var1.pred
all.equal(z, z.hat)
```

Again, the output of Inverse Distance Weighting prediction *is* the observed values. 

  * The RMSE for the transmissivity data using Inverse Distance Weighting is  
    
```{r Caclulate IDW predictions, include=FALSE}
loocv_idw_seq <- seq(1:40)
cv.idw <- function(x, sdata,p){
  if(class(sdata) != "SpatialPointsDataFrame"){
    sp::coordinates(sdata) = ~x+y
  }
  t <- sdata[-x,]
  vg <- gstat::variogram(logT~1,data=t)
  fit.vg <- gstat::fit.variogram(vg, initial, fit.method=2)
  zhat_idw <- idw(formula = logT~1,locations = t,newdata=sdata[x,],idp=p,debug.level=0)
  return(zhat_idw$var1.pred)
}
loocv_idw_p1_resids <- wipp$logT - sapply(loocv_idw_seq,FUN = cv.idw,sdata=wipp,p=1)
loocv_idw_p2_resids <- wipp$logT - sapply(loocv_idw_seq,FUN = cv.idw,sdata=wipp,p=2)
rmse.cv.idw1 <- sqrt(mean(loocv_idw_p1_resids^2))
rmse.cv.idw2 <- sqrt(mean(loocv_idw_p2_resids^2))
```

```{r Return IDW RMSEs, echo=FALSE}
as.data.frame(cbind(rmse.cv.idw1, rmse.cv.idw2)) %>% rename(p1=rmse.cv.idw1, p2=rmse.cv.idw2)
```
 

  * Plot the predicted vs. observed values. Identify the worst prediction 
    
```{r Worst IDW Prediction Plot, echo=FALSE}
par(mfrow=c(1,2))
quilt.plot(x = as.matrix(as.data.frame(wipp)), y=loocv_idw_p1_resids,main="Residuals of IDW (p=1)\n With Worst Prediction Circled",xlab="x",ylab="y")
points(wipp[which.max(loocv_idw_p1_resids),1],wipp[which.max(loocv_idw_p1_resids),2],pch=1,cex=5)

quilt.plot(x = as.matrix(as.data.frame(wipp)), y=loocv_idw_p2_resids,main="Residuals of IDW (p=2)\n With Worst Prediction Circled",xlab="x")
points(wipp[which.max(loocv_idw_p2_resids),1],wipp[which.max(loocv_idw_p2_resids),2],pch=1,cex=5)

```

As with the kriging prediction, observation 35 is predicted the worst. 

  * Compare the RMSE of IDW and Kriging predictions
```{r, echo=FALSE}
as.data.frame(cbind(rmse.cv,rmse.cv.idw1, rmse.cv.idw2)) %>% rename(Krige = rmse.cv,IDWp1=rmse.cv.idw1, IDWp2=rmse.cv.idw2)
```

Predicting using Kriging based on the spherical model fit is better than using Inverse Distance Weighting.  